---
title: "NYCFLIGHTS AND BABY NAMES"
output: md_document
---

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

# install.packages("nycflights13")
library(nycflights13)

#Using Nycflights dataframe we are analyzing wind speed throughout the 2013 year.
nycflights13::weather
coll <- c(1:4, 10:11)
wind <- nycflights13::weather[,coll]

#Plot frame to view any outliers, there seems to be an outlier
ggplot(wind, aes(x=month, y=wind_speed, na.rm=TRUE))+
  geom_point()+
  labs(x = "Month", y = "Wind Speed")


#Get rid of the outlier wind speed
wind <- wind %>% 
  filter(wind_speed < 250)

#plot wind speed by month without the outlier
ggplot(wind, aes(x=month, y=wind_speed, group=month))+
  geom_point()+
  labs(x= "Month", y="Wind Speed")

#average wind speed per month for 2013
wind %>%
  group_by(month) %>%
  summarise(avg.wind = mean(wind_speed)) %>%
  ggplot(aes(x=month, y=avg.wind, group = month)) +
    geom_point() +
    labs(x="Month", y="Wind Speed")
```
The second part of using weather in nycflights is to see the meadian speed at each airport. Make a table and a plot of median wind speed by direction, for each airport. (Can use coord_polar())


```{r}

library(dplyr)
library(tidyr)
library(ggplot2)

#Filter out the NA values for wind direction and use the columns that will be of use.
nycflights13::weather
cols <- c(1:4, 9:10)
direction <- nycflights13::weather[,cols]
valuedirection <- filter(direction, !is.na(wind_dir))

#grouping by the airport and the wind direction and calculating the med wind speed by origin.
directiongroup <- valuedirection %>%
  group_by(origin, wind_dir) %>%
  summarize(median.speed = median(wind_speed)) %>%
  .[order(.$origin),]
  
 #final plot of wind direction by wind speed and origin.
 ggplot(directiongroup, aes(x=wind_dir, y=median.speed, group=origin)) +
  geom_point()+
   facet_wrap(~origin)+
   labs(x="Wind Direction", y="Median Wind Speed")

```

Hurray !! Now we use nycflights13::flights and nycflights13::airlines:
Make a table with two columns: airline name (not carrier code) and median distance flown from JFK airport. The table should be arranged in order of decreasing mean flight distance. Hint: use a _join function to join flights and airlines.
Make a wide-format data frame that displays the number of flights that leave Newark ("EWR") airport each month, from each airline

```{r}

library(dplyr)
library(tidyr)
library(ggplot2)

col <- c(10, 13:14, 16)
flightsdist <- nycflights13::flights[,col] %>%
 filter(., origin == "JFK") %>%
  group_by(carrier)%>%
  summarize(med.dist = median(distance))

total.med.dist.carrier <- left_join(nycflights13::airlines, flightsdist, by="carrier") %>%
  filter(., !is.na(med.dist))

#Final Table for the median distance for each carrier leaving JFK
final.dis.table <- total.med.dist.carrier[,2:3]

```
Make a wide-format data frame that displays the number of flights that leave Newark ("EWR") airport each month, from each airline

```{r}

#Want origin to be EWR, need month and carrier.

col <- c(2, 10, 13)
flightsEWR <- nycflights13::flights[,col] %>%
 filter(., origin == "EWR") %>%
  group_by(carrier, month) %>%
  summarize(Number = n()) %>%
  spread(month, Number)
         

```

Now we move on to babynames!
Identify the ten most common male and female names in 2014. Make a plot of their frequency (prop) since 1880. (This may require two separate piped statements).
Make a single table of the 26th through 29th most common girls names in the year 1896, 1942, and 2016
Write task that involves some of the functions on the Data Wrangling Cheat Sheet and execute it. You may either use your own data or data packages (e.g., the ones listed here).


```{r}
#install.packages("babynames")

babynames::babynames 


Girls <- babynames::babynames %>%
  filter(sex == "F", year == "2014")%>%
  top_n(10)%>%
  select(name) %>%
  left_join(babynames::babynames, by = "name") %>%
  filter(sex == "F")%>%
  ggplot(aes(x=year, y=prop))+
  geom_point()+
  facet_wrap("name")+
  labs(x="Year", y="Proportion")
  
print(Girls)

  
  
Boys <- babynames::babynames %>%
  filter(sex == "M", year == "2014")%>%
  top_n(10)%>%
  select(name) %>%
  left_join(babynames::babynames, by = "name")%>%
  filter(sex == "M")%>%
  ggplot(aes(x=year, y=prop))+
  geom_point()+
  facet_wrap("name")+
  labs(x="Year", y="Proportion")

  print(Boys)
  

  #Make a single table of the 26th through 29th most common girls names in the year 1896, 1942, and 2016 (Except there isnt 2016 data ....)
  
  Girl26.29 <- babynames::babynames %>%
    filter(sex=="F", year=="1896"|year=="1942"|year=="2016")%>%
    group_by(year)%>%
   top_n(29, prop)  %>%
   top_n(-4, prop) %>%
    print()
```







